---
title: "Tidy Tuesday"
author: "Daniel Molitor"
date: "7/14/2020"
output: html_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```
This week's data on astronauts can be found [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-14/readme.md)!

Time to load packages and read in the data. This week I wanted to explore choropleth maps, so I didn't do too much data exploration and mostly focused
on creating a choropleth map using a couple different methods.
```{r}
library(choroplethr)
library(choroplethrMaps)
library(ggmap)
library(ggplot2)
library(magrittr)
library(maps)
library(RColorBrewer)
library(tidyverse)

astronauts <- read_csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-14/astronauts.csv"
)
```

For `choroplethr` and `ggplot2` versions of choropleth maps, the country names need to be correctly formatted. The next step is to format the names accordingly.
In this chunk, `nationality_chor` is the name for `choroplethr` and `nationality_map` is for `ggplot2`.
```{r}
astronauts %<>%
  mutate(
    nationality       = case_when(
      nationality == "Republic of South Africa" ~ "South Africa",
      nationality == "U.K./U.S."                ~ "United Kingdom",
      nationality == "U.S.S.R/Russia"           ~ "Russia",
      nationality == "U.S.S.R/Ukraine"          ~ "Ukraine",
      nationality == "UAE"                      ~ "United Arab Emirates",
      nationality == "U.S."                     ~ "United States of America",
      nationality == "U.K."                     ~ "United Kingdom",
      nationality == "Czechoslovakia"           ~ "Czech Republic",
      nationality == "Hungry"                   ~ "Hungary",
      nationality == "Korea"                    ~ "South Korea",
      nationality == "Malysia"                  ~ "Malaysia",
      nationality == "Netherland"               ~ "Netherlands",
      TRUE                                      ~ nationality
    ),
    nationality_chor = tolower(nationality),
    nationality_map   = case_when(
      nationality == "United Kingdom"           ~ "UK",
      nationality == "United States of America" ~ "USA",
      TRUE                                      ~ nationality
    )
  )
```

The next step is to count the number of distinct astronauts per country. I also need to load the map data that is associated with `choroplethr`,
specifically because I don't want Antarctica to show up in the map.
```{r}
country_count <- astronauts %>%
  distinct(nationality_chor, 
           name, 
           .keep_all = TRUE) %>%
  count(nationality_chor) %>%
  rename(region = nationality_chor,
         value  = n)

# Load map data
data("country.map")
```

Now I want to use `choroplethr` to plot a choropleth where the country fill colors are determined by the count of astronauts originating in the 
respective countries. Since `choroplethr` returns a `ggplot` object, I can tag on any standard `ggplot2` formatting options.
```{r}
country_choropleth(country_count,
                   num_colors = 8,
                   zoom = sort(unique(country.map$region))[-5],
                   title = "Astronauts by Nationality") +
  theme_void() +
  theme(panel.grid = element_blank(),
        legend.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank()) +
  labs(x = "",
       y = "") +
  scale_fill_brewer(palette = "YlOrRd") +
  theme(legend.title = element_blank())
```

Since the country names needed for `choroplethr` are slightly different from those needed by `ggplot2`, I need to retabulate the number of distinct
astronauts in each country, this time using the `ggplot2` version of country name.
```{r}
country_count <- astronauts %>%
  distinct(nationality_map, 
           name, 
           .keep_all = TRUE) %>%
  count(nationality_map) %>%
  rename(region = nationality_map,
         value  = n) %>%
  full_join(map_data("world"),
            by = "region")
```

Now I want to recreate the same (essentially) choropleth map, this time using `ggplot2`. First, a few notes. Since there are a bunch of countries with
a small number of astronauts and a couple with a ton (U.S. and Russia in particular), using the default color gradient will end up masking the subtle
differences on the lower end of the scale. To handle this I manually defined the color scale breaks within the `scale_fill_distiller` call using
`values = c(0, 0.005, .03, ..1, .15, .25, .4, .6, .8, 1)`. From what I can tell, `choroplethr` doesn't allow you to manually override the default scale.
Also, `ggplot2` allows me to change the `NA` color where `choroplethr` doesn't allow that flexibility from what I can tell.
```{r}
ggplot(country_count) +
  geom_polygon(aes(x = long, 
                   y = lat,
                   group = group,
                   fill = value)) +
  coord_map(xlim = c(-180, 180),
            ylim = c(-50, 90)) +
  scale_fill_distiller(
    palette = "YlOrRd",
    direction = 1,
    values = c(0, .005, .015, .03, .1, .15, .25, .4, .6, .8, 1),
    na.value = "black"
  ) +
  theme_dark() +
  theme(panel.grid = element_blank(),
        legend.title = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank()) +
  labs(x = "",
       y = "",
       title = "Astronauts by Nationality")
```

Although `choroplethr` seems to be a pretty nice package, the labels look funky and it seems to have much less flexibility than `ggplot2`.
I, at least, will be sticking with `ggplot2` for any future choropleth needs.

Thanks for reading all the way to the end of this boring post!

[Daniel](https://dmolitor.com)