---
title: "Tidy Tuesday"
author: "Daniel Molitor"
date: "8/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

We back. This week's Avatar: The Last Airbender data can be found [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-08-11/readme.md).
This week's data lent itself to word clouds, so that's exactly what I did.
I relied solely on `wordcloud2` for this venture. Per the usual, the first step
is to load the packages and data.
```{r}
library(tidyverse)
library(wordcloud2)

avatar <- read_csv(
  'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-08-11/avatar.csv'
)
```

The next step is to pull strip all punctuation and digit characters from
the `full_text` field along with other random characters.
```{r}
avatar <- avatar %>%
  mutate(full_text = str_replace_all(full_text,
                                     "[\\[\\]!.?,;'\"\\(\\)0-9]", 
                                     ""))
```

After cleaning the text field, I want to split the text field into distinct
words and count the frequency of each word. I will also remove a bunch of
words that I think are either boring (aka uninformative) or nonsensical.
**Note**: I name the columns in my tibble `word` and `freq`, as `wordcloud2`
requires this naming convention.
```{r}
words <- tibble(
  word = unlist(str_split(avatar$full_text, pattern = " "))
) %>%
  mutate(
    word = trimws(tolower(word)),
  ) %>%
  count(word) %>%
  rename(freq = n) %>%
  filter(!word %in% c("a", 
                      "and",
                      "but",
                      "or",
                      "the",
                      "of",
                      "to",
                      "you",
                      "me",
                      "she",
                      "her",
                      "hers",
                      "he",
                      "his",
                      "him",
                      "at",
                      "as",
                      "is",
                      "it",
                      "in",
                      "i",
                      "on",
                      "",
                      "\t",
                      "-",
                      "--")) %>%
  arrange(desc(freq))
```

Now to make the actual word clouds. `wordcloud2` functionality supports
wordclouds in the shape of an imported image. I'm only showing a vanilla
wordcloud because `wordcloud2` has some serious issues when plotting
using a background image.
```{r  dpi = 150}
wordcloud2(words,
           size = 1,
           backgroundColor = "#282a36",
           color = rep(c("#ff79c6",
                         "#8be9fd",
                         "#f8f8f2",
                         "#bd93f9",
                         "#6272a4"),
                       nrow(words)),
           shuffle = FALSE,
           rotateRatio = 0.25)
```

And that's it! Although `wordcloud2` has some cool features, it also has its
share of problems. In particular, when trying to incorporate a wordcloud
over a background image or use the `letterCloud` function, it either doesn't
work at all or, as in my case, the code needs to be re-run several times
before it works. [This Github issue](https://github.com/Lchiffon/wordcloud2/issues/12)
outlines these problems in more detail. However, the vanilla wordcloud
implementation works seemlessly and overall is really nice! So, for standard
wordclouds I wouldn't hesitate to recommend `wordcloud2`, but for more
advanced features I'd probably turn elsewhere.

Another day, another $

[Daniel](https://dmolitor.com)