---
title: "Tidy Tuesday"
author: "Daniel Molitor"
date: "7/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

This week's penguin data can be found [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-07-28/readme.md)! The first step is to load the data and packages. I wanted to learn how to do a radar plot this week, so I'm focusing on using `plotly` to generate the desired plot.
```{r}
library(plotly)
library(tidyverse)

penguins_raw <- read_csv(
  'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-28/penguins_raw.csv'
)
```

I want to use the radar plot to take a look at the penguins' sizes, split out by the three islands that are represented in the data.
I'll begin by summarizing the average weight, bill depth, bill length, and flipper length of the penguins on each island.
```{r}
measures <- penguins_raw %>%
  mutate(`Body Mass (g)` = as.numeric(`Body Mass (g)`)/1000,
         `Flipper Length (mm)` = as.numeric(`Flipper Length (mm)`)/100,
         `Culmen Length (mm)` = as.numeric(`Culmen Length (mm)`)/10,
         `Culmen Depth (mm)` = as.numeric(`Culmen Depth (mm)`)/10) %>%
  group_by(Island) %>%
  summarise("Body Mass\n(kg)" = mean(`Body Mass (g)`, na.rm = TRUE),
            "Bill Length (cm)" = mean(`Culmen Length (mm)`, na.rm = TRUE),
            "Bill Depth\n(cm)" = mean(`Culmen Depth (mm)`, na.rm = TRUE),
            "Flipper Length (cm/10)"= mean(`Flipper Length (mm)`, na.rm = TRUE)) %>%
  pivot_longer(cols = `Body Mass\n(kg)`:`Flipper Length (cm/10)`,
               names_to = "measure",
               values_to = "value")
```

Now to use a radar plot to look at these four size dimensions by island. As you can see, `plotly` has some very cool, built-in interactivity that can be played with here.
```{r dpi = 150}
plot_ly(
  type = "scatterpolar",
  mode = "lines",
  fill = "toself"
) %>%
  add_trace(
    r = measures %>% filter(Island == "Biscoe") %>% pull(value),
    theta = measures %>% filter(Island == "Biscoe") %>% pull(measure),
    name = "Biscoe",
    fillcolor = "#ff79c6",
    opacity = .8,
    line = list(
      color = "#ff79c6"
    )
  ) %>%
  add_trace(
    r = measures %>% filter(Island == "Dream") %>% pull(value),
    theta = measures %>% filter(Island == "Dream") %>% pull(measure),
    name = "Dream",
    fillcolor = "#8be9fd",
    opacity = .8,
    line = list(
      color = "#8be9fd"
    )
  ) %>%
  add_trace(
    r = measures %>% filter(Island == "Torgersen") %>% pull(value),
    theta = measures %>% filter(Island == "Torgersen") %>% pull(measure),
    name = "Torgersen",
    fillcolor = "#bd93f9",
    opacity = .8,
    line = list(
      color = "#bd93f9"
    )
  ) %>%
  layout(
    title = list(
      text = "Penguin Size",
      size = 3,
      color = "#bd93f9",
      x = 0.012,
      y = 0.988
    ),
    polar = list(
      radialaxis = list(
        visible = TRUE,
        range = c(0, 5),
        showgrid = TRUE,
        showline = FALSE,
        color = "#f8f8f2",
        tickcolor = "#f8f8f2",
        tickvals = c(0, 1, 2, 3, 4, 5)
      ),
      bgcolor = "#282a36"
    ),
    paper_bgcolor  = "#282a36",
    font = list(
      color = "#f8f8f2"
    )
  )
```

My main takeaway is that `plotly` seems to have some nice templates for what would otherwise be pretty difficult-to-create plots using a package such as `ggplot2`. I'm not switching to it for the bulk of visualizations, but I like it a lot!

Until next time,

[Daniel](https://dmolitor.com)