---
title: "Tidy Tuesday"
author: "Daniel Molitor"
date: "8/25/2020"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

I'm back. I know I was desparately missed in my one week of absence, 
but no fear; I'm here! This week's Chopped data can be found
[here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-08-25/readme.md)!
I mostly constructed this on a plane so I didn't delve into anything
too new or original. Really the only completely new thing this time around
was the use of annotations and arrows, which is actually really helpful.

As always, the first step is to load packages and data.
```{r}
library(ggplot2)
library(tidyverse)
library(viridis)

chopped <- read_tsv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-08-25/chopped.tsv"
)
```

Now I want to remove all observations that are missing an IMDB rating.
I also pivot the data longer to easily summarize ratings by judge and fix one
clearly messed up judge's name.
```{r}
chopped <- chopped %>%
  filter(!is.na(episode_rating)) %>%
  pivot_longer(judge1:judge3,
               names_to = "judge_no",
               values_to = "judge") %>%
  mutate(judge = str_replace_all(judge, 
                                 "Aarón Sanchez",
                                 "Aarón Sánchez"))

# Calculate average IMDB scores for individual judges
judge_ratings <- chopped %>% 
  group_by(judge) %>% 
  summarise(mean_rating = mean(episode_rating, na.rm = T),
            n_ep = n()) %>% 
  arrange(desc(mean_rating)) %>%
  rename("Number of\nepisodes" = n_ep)
```

Next, I need to calculate the weighted average of IMDB ratings across
all judges.
```{r}
wa <- sum(judge_ratings$mean_rating*judge_ratings$`Number of\nepisodes`)/
  sum(judge_ratings$`Number of\nepisodes`)
```

Finally, to make the plot. It's ordered from highest to lowest average
IMBD rating, and the bars are colored by the number of episodes each judge
participated in. I've added a couple notations and arrows, which was
new territory for me, but will be extremely useful for future plots.
```{r dpi=300}
ggplot(judge_ratings) + 
  geom_hline(yintercept = wa, 
             color = "white", 
             linetype = "dashed") + 
  geom_bar(aes(x = factor(judge, levels = rev(unique(judge)), ordered = T), 
               y = mean_rating, 
               fill = `Number of\nepisodes`), 
           stat = "identity", 
           width = 1) + 
  geom_curve(aes(x = judge[83], 
                 y = wa + .5, 
                 xend = judge[83], 
                 yend = wa + .05), 
             color = "white",
             curvature = -.2, 
             arrow = arrow(angle = 45, 
                           length = unit(3, units = "mm"))) + 
  geom_curve(aes(x = judge[13], 
                 y = wa + .8, 
                 xend = judge[10], 
                 yend = wa + .5), 
             color = "white",
             curvature = .4, 
             arrow = arrow(angle = 45, 
                           length = unit(3, units = "mm"))) +
  annotate("text", 
           label = "Average IMDB rating\nfor all judges (8.38)", 
           color = "white",
           fontface = "bold",
           x = judge_ratings$judge[80], 
           wa + .8) + 
  annotate("text", 
           label = "A bunch of\none-hit wonders!", 
           color = "white", 
           fontface = "bold",
           x = judge_ratings$judge[15], 
           wa + .9) +
  scale_fill_viridis(option = "E", 
                     direction = 1) + 
  coord_flip() + 
  theme(panel.grid = element_blank(), 
        plot.background = element_rect(color = "#44475a", 
                                       fill = "#44475a"), 
        panel.background = element_rect(color = "#44475a", 
                                        fill = "#44475a"), 
        legend.background = element_rect(color = "#44475a", 
                                         fill = "#44475a"), 
        legend.text = element_text(color = "white",
                                   face = "bold"), 
        legend.title = element_text(color = "white",
                                    face = "bold"), 
        axis.text = element_text(color = "white",
                                 size = 8),
        axis.title = element_text(color = "white",
                                  face = "bold"),
        plot.title = element_text(color = "white",
                                  hjust = 0.5,
                                  face = "bold")) +
  labs(title = "Average IMDB Ratings for Judges",
       y = "IMDB Rating",
       x = "")
```

As always, the image looks like poop in R Markdown. When running the code in
RStudio you can expand the plot window to full screen and it looks great.
With that unimpressive outro, we've concluded this week's exploration.

Until next time!

[Daniel](https://dmolitor.com)